---
title: "Week 9, Day 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(PPBDS.data)
library(rstanarm)
library(tidyverse)

# We will use the `kenya` dataset from PPBDS.data. This dataset is from Harris,
# Kamindo and Van der Windt (2020): "Electoral Administration in Fledgling
# Democracies: Experimental Evidence from Kenya." The authors worked with
# Kenya's electoral commission in 1,674 communities by assigning polling
# stations to either a control group or to one of five other treatments.

# For simplicity we will only consider polling stations assigned to either
# "control" or to "local". The latter means a treatment in which election
# officials visited the community's polling station for two days, thereby making
# it easier for people to register.

# The outcome variable, `reg_chg`, is the change in voter registration in each
# community. `poverty` is the percentage of the community below the poverty
# level. `distance` is the number of kilometers between the polling station and
# the central political office (where registrations normally take place).
# `pop_density` is a measure of population density around the polling station.


week_9 <- kenya %>% 
  rename(reg_chg = reg_byrv13) %>% 
  filter(treatment %in% c("control", "local")) %>% 
  droplevels() %>% 
  select(reg_chg, treatment, poverty, 
         distance, pop_density)
```


## Scene 1

**Prompt:** In addition to the outcome variable and the treatment, we have 3 statistics for the communities in the sample --- poverty, distance and population density. See code comments above for details. Never hurts to do a little exploratory data analysis. Look at the data!

* Make a scatter plot of `poverty` on the y-axis and `distance` in the x-axis. Include a `geom_smooth()` with `method` set to "lm" and `se` equal to FALSE. Write a sentence to answer these questions:

```{r}
week_9 %>%
  ggplot(aes(distance, poverty)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```


  + Does poverty cause distance? Or does distance cause poverty? Or both? Or neither?
  
not a causation, correlation

  + How accurate is the model beyond 90 kilometers?  
  
not as accurate, because there aren't many data points
  
  + Eye-balling the chart, how different is the poverty rate between two polling stations whose distance differs by 50 kilometers?
  
.48 poverty -.4 = .8 difference, 80% dif
  
  
* Fit a `stan_glm()` model which relates poverty to distance. Print a summary. (Hint: In order to see the coefficient of `distance`, you will probably need to use the `digits` argument to `print()`.) Interpret the meaning of the two important parameters.  

```{r}
fit_model <- stan_glm(poverty ~ distance,
                      data = week_9,
                      refresh = 0)

print(fit_model, digit = 4)
```
intercept median is the average median of the posterior prob distribution of poverty when distance = 0


distance value* km value and add to control to get poverty level per difference



## Scene 2

**Prompt:** Let's make a model with `reg_chg` as the outcome variable and with treatment, poverty, distance and pop_density as the predictors. (Using outcome/predictors is the typical nomenclature in statistics. In Economics, we would refer to dependent/independent variables.)

* Provide the math for this model.

$$ reg\_chg_i = \beta_{int} + \beta_1 x_{treated,i} + \beta_2 x_{poverty,i} + \beta_3 x_{distance,i} + \beta_4 x_{pop\_density,i} + \epsilon_i$$
i = observation

* Estimate the model. Assign the fitted model to `fit_2`. Interpret the important parameters of the model.

```{r}
fit_2 <- stan_glm(data = week_9,
                  reg_chg ~ treatment + poverty + distance + pop_density,
                  refresh = 0)

print(fit_2, digit = 6)

```
intercept: reg_chg is when the predictors are 0
treatmentlocal: to get the reg_chg values you would just add treatmentlocal to intercept
poverty: 
distance: you would multiple the distance beta value by the distance and add it to the intercept value to get the reg_chg for that point
pop_density

this only works when it is a line; 



## Scene 3

**Prompt:** Let's consider a different model. First, create a new variable, `poverty_n`, which is just `poverty` minus the mean of `poverty` then divided by the standard deviation of `poverty`. The "n" is for "normalization."  Second, consider `reg_chg` as a function of `treatment`, `poverty_n` and the interaction between the two.

```{r}
week_9 <- week_9 %>% mutate(poverty_n = (poverty - mean(poverty))/sd(poverty))
```


* Fit the  model and interpret the coefficients. Focus on the coefficient of the interaction term.

```{r}
fit_3 <- stan_glm(reg_chg ~ treatment + poverty_n,
                  data = week_9,
                  refresh = 0)
```


* Create a graphic of the posterior probability distribution of the causal effect of `local` treatment in polling stations which have +2 sigma poverty rates. Note that there are two ways to do this. First, we can use a parameter approach. This is what we did in class last week. Second, we can use `posterior_epred()`, which if generally easier since it does the math for us. Confirm that the two approaches give the same answer.






